---
title: "Downstream analysis of phyloacc - nectar without dollo - CNEE set definition - follow up2"
author: "maggie"
date: "2024-08-19"
output:
  html_document: 
    toc: yes
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
# This is a way to set options for all code chunks at once
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, setup=TRUE, include=FALSE}
.libPaths()
libpath = "/opt/R/4.2.0/lib64/R/library"
# # library(tidyverse, lib.loc = libpath) # 2.0.0
# library(tidytree) # 0.4.6
# library(ggtree, lib.loc = libpath)
library(tidyverse)
library(treeio)
library(ggtree)
# devtools::install_github("YuLab-SMU/ggmsa", lib.loc = libpath) 
library(ggmsa)
library(Biostrings)

theme_m = theme_classic() + theme(legend.key.size = unit(8, 'pt'))
```

## load ids for conversion
```{r, include=FALSE, cache=TRUE, cache.extra = tools::md5sum('/home/mpg08/mko/Nectar/analysis/CNEEanalysis_2021/01b_cnee_prep/bed_outputs/galGal6_final_conserved_CNEEs.bed'), cache.extra = tools::md5sum( '/home/mpg08/mko/Nectar/analysis/CNEEanalysis_2021/03_phyloacc/2024ver/inputs/mod/galloseq.part.bed')}

# for getting coordinates
tmp_path = paste0("/home/mpg08/mko/Nectar/analysis/CNEEanalysis_2021/01b_cnee_prep/bed_outputs/galGal6_final_conserved_CNEEs.bed")
ggal_final_cnees = read_tsv(tmp_path, col_names = c("chr", 'start', 'end', 'id'))
nrow(ggal_final_cnees) # 363747

# for getting original cnee ID
tmp_path = paste0("/home/mpg08/mko/Nectar/analysis/CNEEanalysis_2021/03_phyloacc/2024ver/inputs/mod/galloseq.part.bed")
galloseq.part = read_tsv(tmp_path, col_names = c('id', 'start', 'end'))
nrow(galloseq.part) # 363476

```

## load phyoacc outputs
```{r load_phyloacc_outputs, include=FALSE, cache=TRUE, cache.extra = tools::md5sum('/home/mpg08/mko/Nectar/analysis/CNEEanalysis_2021/03_phyloacc/2024ver/nectar_nodollo/phyloacc-out-02-07-2024.11-55-54/results/elem_lik.txt')}
project_path = paste0(getwd(), "/phyloacc-out-02-07-2024.11-55-54/results/")
# project_path = paste0(getwd(), "/CNEEanalysis_2021/03_phyloacc/2024ver/nectar_nodollo/phyloacc-out-02-07-2024.11-55-54/results/")

# 1.  score
path = paste0(project_path, "elem_lik.txt")
score = read_tsv(path, comment = "#")
dim(score) # 363213     21
# final set total after aligned: 363476
sum(is.na(score)) # 63156
sum(is.na(score$accel.lineages.m1)) # 56284
sum(is.na(score$accel.lineages.m2)) # 4184
sum(is.na(score$conserved.lineages.m1)) # 336
sum(is.na(score$conserved.lineages.m2)) # 336

# 2. posterior Z of M2
path = paste0(project_path, "rate_postZ_M2.txt")
postZ = read_tsv(path) # %>% filter(!is.na(n_rate))
dim(postZ) # 362877    354
sum(is.na(postZ$n_rate)) # 0
sum(is.na(postZ)) # 0

postZ_6 = postZ[, 1:6] 

# full model post. probability
ind = grep("_3", names(postZ))
# postZp = score %>% dplyr::select(No., ID) %>%
#   right_join(postZ %>% dplyr::select(all_of(c(1, ind))) ) # pp of acc. CNEE
postZp = postZ %>% dplyr::select(all_of(c(1, ind)))  # pp of acc. CNEE
dim(postZp) # 362877     88
sum(is.na(postZp$`Locus ID`)) # 0

# x = postZp %>% filter(is.na(ID)) # for checking
# postZp = postZp %>% filter(!is.na(ID))
names(postZp) =  gsub("_3", "", names(postZp))

# missing species post. probability
ind = grep("_0", names(postZ))
postZm = postZ %>% dplyr::select(all_of(c(1, ind))) # pp of acc. CNEE
dim(postZm) # 362877     88
names(postZm) =  gsub("_0", "", names(postZm))

# for id conversion
score_ids = score %>%
  mutate(id = galloseq.part$id[original.id]) %>%
  # filter(logbf1 > 0 & logbf2 > 0 & logbf3 > 0) %>%
  left_join(postZ_6, by = c("phyloacc.id" = "Locus ID")) %>%
  bind_cols(ori = 'nectar_nodollo')
score_ids_sel = score_ids %>% dplyr::select(matches("id"))


# 3. tree
path = paste0(project_path, "elem_lik.txt")
tree_tx = read_table(path, skip = 35, n_max = 36, col_names = F)
tree = read.tree(text = tree_tx$X3)
plot(tree, cex = 0.5)

# 4. species name table
path = paste0(project_path, "../../input/commonN.txt")
common_name = read_tsv(path, col_names = F)

targets = c("HLlepAsp1", "HLdicExi1", 
            "HLlicPen1", "HLlicMelCas1", "HLphyNov1", "HLgraPic1", 
            "HLlorGal1", "HLtriMol2", "HLaraSol1", "HLamaAes1", 
            "HLphaSup1", "HLfloFus1", "HLcalAnn5") # target species (rm pardalote and cinPul)

# all_sp = treeData$tree$tip.label
# non_target = setdiff(all_sp, targets)

```

## plot functions

```{r, include=FALSE}

plot_single_cnee = function(cneeid, inputdf = postZp, target_species = NULL, show_legend = TRUE, show_all_tips = FALSE, xratio = 0.8){
  # cneeid = '808-17'
  # inputdf = postZp
  # target_species = targets; xratio = 0.8
  
  # pp data
  atmp = inputdf %>% filter(`Locus ID` == cneeid) %>% 
    pivot_longer(cols = 2:last_col(), names_to = 'label', values_to = 'pp')
  
  # missing data
  atmp_m = postZm %>% filter(`Locus ID` == cneeid) %>% 
    pivot_longer(cols = 2:last_col(), names_to = 'label', values_to = 'missing') %>% 
    mutate(
      # group: only show targets and color missing targets grey
      group = ifelse(label %in% target_species, 'Nectar', NA), 
      group = ifelse(missing >0 & group == 'Nectar', 'Missing', group),
      # all_lab: Nectar, non-target, and color missing targets grey; if non-target and missing, this will be a NA
      all_lab = 'Present',
      all_lab = ifelse(missing >0 & group == 'Nectar', 'Missing', all_lab)) 
 
  # get No.
  # aNo = atmp$No.[1] # for postZ # old no need
  
  # get rates and other pp
  atmp_full = postZ %>% filter(`Locus ID` == cneeid)
  ratio1 = atmp_full$c_rate # conserv rate
  ratio2 = atmp_full$n_rate # acc rate
  r1 = round(ratio1, 2)
  r2 = round(ratio2, 2)
  atmp_full_long = atmp_full %>% 
    dplyr::select(matches("_[1|2|3]$")) %>% 
    pivot_longer(cols = matches("_[1|2|3]$"), names_to = c("label", "state"), names_pattern = "(.*)_(\\d)") %>% 
    pivot_wider(id_cols = "label", names_from = "state", values_from = "value") %>% 
    mutate(edge_len = `1` + `2`*ratio1 + `3`*ratio2) %>% 
    dplyr::select(-matches("[1|2|3]"))
  
  # combine
  atmpc = atmp_full_long %>% left_join(atmp) %>% 
    left_join(atmp_m)
  atmpc$group = factor(atmpc$group, levels = c('Nectar', 'Missing', NA))
  atmpc$all_lab = factor(atmpc$all_lab, levels = c('Present', 'Missing', NA))
  
  # get logBFs
  stmp = score %>% filter(`phyloacc.id` == cneeid)
  b1 = round(stmp$logbf1, 2); b2 = round(stmp$logbf2, 2); b3 = round(stmp$logbf3, 2)
  
  # tree + pp data
  atree_for_plot = tree
  # atree_data = treeio::treedata(phylo = atree_for_plot, data = atmpc) # doesn't work
  atree_data = atree_for_plot %>% as_tibble() %>% 
    full_join(atmpc, by = 'label') %>% 
    mutate( branch.length = branch.length*edge_len ) %>% dplyr::select(-edge_len) %>% 
    as.treedata()
  atree_data@phylo$tip.label = tree$tip.label
  # x_limit = max(atree_data@phylo$edge.length) * 0.25
  
  # color scale for pp
  rbPal_tmp = colorspace::diverge_hcl(3, palette = "Blue-Red")
  # high_col = rbPal_tmp[3]; mid_col = rbPal_tmp[2]; low_col = rbPal_tmp[1]
  p_tree = ggtree(atree_data, aes(color = pp)) +
    scale_color_gradientn(colours = rbPal_tmp, 
                          values = scales::rescale(c(0, 0.899999, 1)), 
                          breaks = c(0, signif(0.899999, 2), 1)) +
    # geom_text(aes(label=node), hjust=-.3) +
    # xlim_tree(x_limit) +
    hexpand(xratio) + # larger number, narrower
    ggtitle(cneeid, 
            # subtitle = substitute(paste(r[1], "=", r1, ", ", r[2], "=", r2), list(r1 =round(ratio1,2), r2 = round(ratio2,2)))
            subtitle = bquote(atop("r"[1]~" = "~.(r1)~", r"[2]~" = "~.(r2),
                                       "logbf1 = "~.(b1)~", logbf2 = "~.(b2)~", logbf3 = "~.(b3)))
            ) +
    theme_tree() + 
    theme(plot.title = element_text(hjust = 0.5, size = 8), 
          plot.subtitle = element_text(hjust = 0.5, size = 6),
          legend.text = element_text(size = 6),
          legend.title = element_text(size = 6), 
          legend.key.size = unit(8, 'pt')) ; 
  # p_tree = p_tree %>% ggtree::rotate(44); p_tree # give up for now
  p_tree = p_tree %>% rotate(45); # p_tree # works
  
  if(show_all_tips){
    p_tree = p_tree +
      geom_tiplab(mapping = aes(alpha = all_lab), color = 'black', offset = 0, size = 2) +
      scale_alpha_manual(values = c(1, 0.5), na.value = 1, breaks = c('Present', 'Missing')) +  # here NA is for non-target missing
      guides(alpha = guide_legend(override.aes = aes(label = "-"))) ; p_tree
  }else{
    p_tree = p_tree +
      geom_tiplab(mapping = aes(alpha = group), color = 'black', offset = 0, size = 2) +
      scale_alpha_manual(values = c(1, 0.5), na.value = 0, breaks = c('Nectar', 'Missing')) +
      guides(alpha = guide_legend(override.aes = aes(label = "-"))) ; p_tree
  }
  if(show_legend == F){ p_tree = p_tree + theme(legend.position = 'none') }
  return(p_tree)
}

# lapply('808-17', FUN = plot_single_cnee, target_species = targets) # works
```

## distribution of bayers factors
log Bayes factors thresholds:
-   log-BF1\>20 and log-BF2\>0 (Hu avian)
-   log-BF1\>5 and log-BF2\>5 (Hu mammal)
-   BF1 \>= 10, and BF2 \>= 1 (Sackton)

### distribution of logbf1

BF1 = P(Y\|M1)/P(Y\|M0)

M1: lineage-specific model with specific shift pattern

M0: null model with no acceleration in any lineage

```{r logbf1, fig.height=3, fig.width=10, echo=FALSE}
nbins = 100

# logbf1
x_tmp = score$logbf1
# print(range(x_tmp))
res = hist(x_tmp, breaks = 50, plot = FALSE)
y_anno = max(res$counts, na.rm = T) * 0.6
x_anno = signif(ave(c(median(x_tmp, na.rm = T), max(x_tmp, na.rm = T)))[1] , digits = 2) * 1

pbf1 = ggplot(score, aes(x = logbf1)) +
  geom_histogram(bins = nbins) +
  scale_x_continuous(expand = c(0,0,0.05,0)) +
  scale_y_continuous(expand = c(0,0,0.05,0), name = "Freq") +
  annotate("text", x = x_anno, y = y_anno, 
           label = paste0("median = ", signif(median(x_tmp, na.rm = T), digits = 2), 
                          "\nmax = ", signif(max(x_tmp, na.rm = T), digits = 2), 
                          "\nmin = ", signif(min(x_tmp, na.rm = T), digits = 2)), 
           hjust = 0) +
  geom_vline(xintercept = 0, color = 'red') +
  annotate("text", x = 100, y = 50000, 
           label = "x = 0", 
           hjust = 0.5, color = 'red') +
  theme_m; 

pbf1_2 = ggplot(score, aes(x = logbf1)) +
  geom_histogram(bins = nbins) +
  scale_x_continuous(expand = c(0,0,0.05,0)) +
  scale_y_continuous(expand = c(0,0,0.05,0), name = "Freq") +
  ggtitle('zoom in') +
  coord_cartesian(xlim = c(1, 100)) +
  theme_m + theme(plot.title = element_text(hjust = 0.5)); 

cowplot::plot_grid(pbf1, pbf1_2, nrow = 1)

```

### distribution of logbf2

BF2 = P(Y\|M1)/P(Y\|M2)

M1: lineage-specific model with specific shift pattern

M2: full model allowing arbitrary shifts

```{r logbf2, fig.height=3, fig.width=10, echo=FALSE}
nbins = 100

# logbf2
x_tmp = score$logbf2
# print(range(x_tmp))
res = hist(x_tmp, breaks = 50, plot = FALSE)
y_anno = max(res$counts, na.rm = T) * 0.6
# x_anno = signif(median(x_tmp, na.rm = T), digits = 2) * -1*50
x_anno = signif(ave(c(median(x_tmp, na.rm = T), min(x_tmp, na.rm = T)))[1] , digits = 2) * 1

pbf2 = ggplot(score, aes(x = logbf2)) +
  geom_histogram(bins = nbins) +
  scale_x_continuous(expand = c(0,0,0.05,0)) +
  scale_y_continuous(expand = c(0,0,0.05,0), name = "Freq") +
  annotate("text", x = x_anno, y = y_anno, 
           label = paste0("median = ", signif(median(x_tmp, na.rm = T), digits = 2), 
                          "\nmax = ", signif(max(x_tmp, na.rm = T), digits = 2), 
                          "\nmin = ", signif(min(x_tmp, na.rm = T), digits = 2)), 
           hjust = 0) +
  geom_vline(xintercept = 0, color = 'red') +
  annotate("text", x = 10, y = 50000, 
           label = "x = 0", 
           hjust = 0.5, color = 'red') +
  theme_m; 

pbf2_2 = ggplot(score, aes(x = logbf2)) +
  geom_histogram(bins = nbins) +
  scale_x_continuous(expand = c(0,0,0.05,0)) +
  scale_y_continuous(expand = c(0,0,0.05,0), name = "Freq") +
  coord_cartesian(xlim = c(0, 50)) +
  ggtitle('zoom in') +
  theme_m + theme(plot.title = element_text(hjust = 0.5)); 

cowplot::plot_grid(pbf2, pbf2_2, nrow = 1)

```

### distribution of logbf3

BF3 = P(Y\|M2)/P(Y\|M0)

M2: full model allowing arbitrary shifts

M0: null model with no acceleration in any lineage

```{r logbf3, fig.height=3, fig.width=10, echo=FALSE}
nbins = 100

# logbf3
x_tmp = score$logbf3
# print(range(x_tmp))
res = hist(x_tmp, breaks = 50, plot = FALSE)
y_anno = max(res$counts, na.rm = T) * 0.6
x_anno = signif(ave(c(median(x_tmp, na.rm = T), max(x_tmp, na.rm = T)))[1] , digits = 2) * 1

pbf3 = ggplot(score, aes(x = logbf3)) +
  geom_histogram(bins = nbins) +
  scale_x_continuous(expand = c(0,0,0.05,0)) +
  scale_y_continuous(expand = c(0,0,0.05,0), name = "Freq") +
  annotate("text", x = x_anno, y = y_anno, 
           label = paste0("median = ", signif(median(x_tmp, na.rm = T), digits = 2), 
                          "\nmax = ", signif(max(x_tmp, na.rm = T), digits = 2), 
                          "\nmin = ", signif(min(x_tmp, na.rm = T), digits = 2)), 
           hjust = 0) +
  geom_vline(xintercept = 0, color = 'red') +
  annotate("text", x = 100, y = 50000, 
           label = "x = 0", 
           hjust = 0.5, color = 'red') +
  theme_m; 

pbf3_2 = ggplot(score, aes(x = logbf3)) +
  geom_histogram(bins = nbins) +
  scale_x_continuous(expand = c(0,0,0.05,0)) +
  scale_y_continuous(expand = c(0,0,0.05,0), name = "Freq") +
  coord_cartesian(xlim = c(0, 50)) +
  ggtitle('zoom in') +
  theme_m + theme(plot.title = element_text(hjust = 0.5)); 

cowplot::plot_grid(pbf3, pbf3_2, nrow = 1)

```

### distribution of logbf1 to logbf3

```{r  logbf1-3, fig.height=3, fig.width=10, echo=FALSE}
cowplot::plot_grid(pbf1, pbf2, pbf3, nrow = 1)
```

### plotting cness with logbf1 \> 0 & logbf2 \> 0 & logbf3 \> 0 

```{r logbf1-3_gt0, fig.height=3, fig.width=10, echo=FALSE}
cowplot::plot_grid(pbf1_2, pbf2_2, pbf3_2, nrow = 1)
```


### look at CNEEs: prep

```{r prep_bf_cat, fig.height=6, fig.width=4}
score_summary = score %>%
  mutate(bf1_cat_gt0 = ifelse(logbf1 >=0, 1, 0),
         bf1_cat_gt2 = ifelse(logbf1 >=2, 1, 0),
         bf1_cat_gt5 = ifelse(logbf1 >=5, 1, 0),
         bf1_cat_gt10 = ifelse(logbf1 >=10, 1, 0),
         bf1_cat_gt15 = ifelse(logbf1 >=15, 1, 0),
         bf2_cat_gt0 = ifelse(logbf2 >=0, 1, 0),
         bf2_cat_gt2 = ifelse(logbf2 >=2, 1, 0),
         bf2_cat_gt5 = ifelse(logbf2 >=5, 1, 0),
         bf2_cat_gt10 = ifelse(logbf2 >=10, 1, 0),
         bf2_cat_gt15 = ifelse(logbf2 >=15, 1, 0),
         bf3_cat_gt0 =  ifelse(logbf3 >=0, 1, 0),
         bf3_cat_gt2 =  ifelse(logbf3 >=2, 1, 0),
         bf3_cat_gt5 =  ifelse(logbf3 >=5, 1, 0),
         bf3_cat_gt10 =  ifelse(logbf3 >=10, 1, 0),
         bf3_cat_gt15 =  ifelse(logbf3 >=15, 1, 0)) %>% 
  dplyr::select(phyloacc.id, matches("_cat")) 
```

### look at CNEEs: logbf1 \> 10

```{r logbf1_gt_10, fig.height=7, fig.width=10 }
input = score_summary %>% filter(bf1_cat_gt10 == 1)
n_distinct(input$phyloacc.id)
cneeIDs = sample(input$phyloacc.id, 10) # randomly sample n CNEEs to look, each row 3 in
pl = lapply(cneeIDs, FUN = plot_single_cnee, target_species = targets, show_legend = FALSE, show_all_tips = FALSE)
cowplot::plot_grid(plotlist = pl, ncol = 5)
```

### look at CNEEs: 5 \< logbf1 \< 10: some unspecific but fairly ok

```{r logbf1_5_to_10, fig.height=7, fig.width=10 }
input = score %>% filter(logbf1>= 5 & logbf1 <10)
n_distinct(input$phyloacc.id)
cneeIDs = sample(input$phyloacc.id, 10)
pl = lapply(cneeIDs, FUN = plot_single_cnee, target_species = targets, show_legend = FALSE, show_all_tips = FALSE)
cowplot::plot_grid(plotlist = pl, ncol = 5)
```

### look at CNEEs: 5 \< logbf1 \< 10 & logbf2 \> 0:

```{r logbf1_5_to_10_BF2, fig.height=7, fig.width=10 }
input = score %>% filter(logbf1>= 5 & logbf1 <10 & logbf2 > 0 )
n_distinct(input$phyloacc.id)
cneeIDs = sample(input$phyloacc.id, 10)
pl = lapply(cneeIDs, FUN = plot_single_cnee, target_species = targets, show_legend = FALSE, show_all_tips = FALSE)
cowplot::plot_grid(plotlist = pl, ncol = 5)
```

## calculate expectation
### prep: whole tree
```{r calculate_expectation_whole_phylogeny}

tree_tb = tree %>% as_tibble()

get_1deg_parrent = function(branch){
  l = treeio::parent(tree_tb, branch)$label
  out = l
  return(out)
}

postZpl = postZp %>% 
  pivot_longer(cols = 2:last_col(), names_to = 'label', values_to = 'pp') %>% 
  left_join(score_ids_sel, by = c("Locus ID" = "phyloacc.id"))

```

### define clades and function
```{r define_clades}
# def nectar clades ====
#### hummingbird ====
hummingbird_tips = c("HLcalAnn5", "HLfloFus1", "HLphaSup1")
all_parents = unlist(sapply(hummingbird_tips, get_1deg_parrent))
hummingbird_input = unique(c(hummingbird_tips, all_parents)) # 5

#### parrot tips ====
parrot_tips = c("HLlorGal1", "HLtriMol2", "HLaraSol1", "HLamaAes1")
all_parents = unlist(sapply(parrot_tips, get_1deg_parrent))
all_parents2 = unlist(sapply(all_parents, get_1deg_parrent))
parrot_input = unique(c(parrot_tips, all_parents, all_parents2)) # 7

#### honeyeater tips ====
honeyeater_tips = c("HLlicPen1", "HLlicMelCas1", "HLphyNov1", "HLgraPic1")
all_parents = unlist(sapply(honeyeater_tips, get_1deg_parrent))
# pardalote_tips = c("HLparPun1")
# honeyeater_pardalote_input = unique(c(honeyeater_tips, all_parents, pardalote_tips)) # 8
honeyeater_input = unique(c(honeyeater_tips, all_parents)) # 7

#### sunbird+flowerpecker tips ====
sunbird_tips = c("HLlepAsp1", "HLdicExi1")
all_parents = unlist(sapply(sunbird_tips, get_1deg_parrent))
sunbird_input = unique(c(sunbird_tips, all_parents)) # 5 >2024:3

# def non-nectar clades ====
#### swift ====
swift_tips = c("HLapuApu1", "HLchaPel1")
all_parents = unlist(sapply(swift_tips, get_1deg_parrent))
swift_input = unique(c(swift_tips, all_parents)) # 3
#### falcon tips (commented just becasue it's already generated above) ====
falcon_tips = c("HLfalTin1", "falPer1", "falChe1")
all_parents = unlist(sapply(falcon_tips, get_1deg_parrent))
falcon_input = unique(c(falcon_tips, all_parents)) # 5
#### lyrebird ====
lyrebird_tips = c("HLatrCla1", "HLmenNov1")
all_parents = unlist(sapply(lyrebird_tips, get_1deg_parrent))
lyrebird_input = unique(c(lyrebird_tips, all_parents)) # 3
#### passeri -1 ====
passeri1_tips = c("HLtaeGut4", "ficAlb2", "pseHum1")
all_parents = unlist(sapply(passeri1_tips, get_1deg_parrent))
all_parents2 = unlist(sapply(all_parents, get_1deg_parrent))
passeri1_input = unique(c(passeri1_tips, all_parents, all_parents2)) # 8

# def branch class ====
core_nectars = unique(c(hummingbird_input, parrot_input, 
                        honeyeater_input, sunbird_input)) # 25>24>2024:22
reversals = c("HLnymHol2", "HLmalCya1", "HLmalEle1") # 4>3 (20220728 we decide to move "HLempTra1" to ambigous)
ambiguous = unique(c("HLstrHab1", "HLamaAes1-HLstrHab1", "HLamaAes1-HLnymHol2", 
                     "HLparPun1", # pardalote 20220801
              "HLfurRuf1", 
              "HLcliRuf1",
              "HLempTra1", # 0728 addition
              # "HLfurRuf1-HLempTra1", ## Tim thinks it should be in core non-nectar
              "HLmalCya1-HLmalEle1", 
              "HLacaPus1", "HLacaPus1-HLparPun1", "HLlicMelCas1-HLacaPus1",
              "HLparMaj1", "HLmelMel1", "HLserCan1", "HLmelMel1-HLserCan1"))# 13 >14>0728>15
# core_nonnectars = setdiff(tree_tb$label, c(core_nectars, reversals, ambiguous)) # 47
core_nonnectars = unique(c(passeri1_input, lyrebird_input, falcon_input, swift_input)) #19; 20230720 corrected

# sum up the exp (change in pp from the previous branch) of the given branches
clade_specific = function(clade, name_of_clade, thre, plotSingleCNEEs = TRUE){
  # clade = honeyeater_input
  # name_of_clade = 'honeyeaters'
  
  print(name_of_clade)
  
  exp_sel = postZpl %>% 
    # filter(`Locus ID` == '2630-15') %>%
    filter(label %in% clade) %>% 
    group_by(`Locus ID`) %>% 
    # 20240319: i think sum up now is misleading, we actually want 'at least 1 branch > pp_threshol'
    # summarise(exp = sum(pp)) %>%
    summarise(exp = sum(pp>thre)) %>% filter(exp >=1 ) %>% 
    left_join(score %>% dplyr::select(phyloacc.id, matches('logBF')), 
              by = c("Locus ID" = "phyloacc.id")) %>%
    arrange(desc(exp))
  
  # for checking calculation
  # x = postZp %>% dplyr::select(ID, matches(sel))
  
  # tmp_cnee = exp_sel$id[exp_sel$exp>0]
  # x = exp_sel %>% filter(exp > 0) %>% arrange(desc(exp))
  # tmp_cnee = x$id[1:8]
  
  sum(exp_sel$exp>0.9) # 32279 > 20653
  sum(exp_sel$exp>0.95) # 26668 > 17396
  sum(exp_sel$exp>1) # 17065 > 13929
  
  # thre = 1
  nbins = 25
  
  # clade
  y_max = 300000
  n = sum(exp_sel$exp >= thre, na.rm = T); n
  p_exp_hum = ggplot(exp_sel, aes(x = exp)) +
    geom_histogram(bins = nbins, fill = 'steelblue') +
    scale_x_continuous(expand = c(0,0,0.05,0)) +
    scale_y_continuous(expand = c(0,0,0.05,0), limits = c(0, y_max)) +
    ggtitle(paste0(name_of_clade, " (", n_distinct(clade), ")\n", 
                   "Num (expectation >= ", thre, "): ", n)) +
    theme_m; p_exp_hum
  
  # clade > 0.95
  y_max2 = 15000
  n = sum(exp_sel$exp >= thre, na.rm = T); n
  p_exp_hum95 = ggplot(exp_sel %>% filter(exp >= thre), aes(x = exp)) +
    geom_histogram(bins = nbins, fill = 'steelblue') +
    scale_x_continuous(expand = c(0,0,0.05,0)) +
    scale_y_continuous(expand = c(0,0,0.05,0), limits = c(0, y_max2)) +
    ggtitle(paste0(name_of_clade, " (", n_distinct(clade), ")\n", 
                   "Num (expectation >= ", thre, "): ", n)) +
    theme_m; p_exp_hum95
  
  # logbf3 vs exp hex
  nbins = 25
  corv = signif(cor(exp_sel$exp, exp_sel$logbf3, "complete.obs"), 2); corv # 0.29
  p_hum_bf3_hex = ggplot(exp_sel, aes(x = exp, y = logbf3)) + # , color = interest
    geom_hex(bins = nbins) + 
    scale_fill_distiller("Blues", direction = 1) +
    scale_x_continuous(expand = c(0,0,0.05,0)) +
    scale_y_continuous(expand = c(0,0,0.05,0)) +
    ggtitle(name_of_clade) +
    theme_m + theme(legend.position = 'right'); p_hum_bf3_hex
  
  if(plotSingleCNEEs){ 
    # plot single cnee - top5 exp
    ind = 1:5
    tmp_cnee = exp_sel$`Locus ID`[ind]
    tmp_mean = exp_sel$exp[ind]
    input_titles = paste0(", exp_mean=", signif(tmp_mean, 2))
    
    ind_in_score = sapply(tmp_cnee, function(s){which(score$phyloacc.id == s)})
    if(!identical(tmp_cnee,  score$phyloacc.id[ind_in_score])){
      print('checking cnee ordering [plot single cnee')}
    
    Ncol = 5
    Nrow = ceiling(length(ind_in_score)/Ncol)
    }
  
  return(list(exp_sel, p_exp_hum, p_exp_hum95, p_hum_bf3_hex))
}
```

### expectation output: nectar and non-nectar:
```{r cal_nectar}
# cal nectar clades (per clade) ====
all_input = list(hummingbird_input, parrot_input,
                 honeyeater_input, sunbird_input)
all_input_names = list("hummingbirds", "parrots",
                       "honeyeaters", "sunbirds")
thre_in = 0.9
cladeouts = mapply(clade_specific,
                 clade = all_input,
                 name_of_clade = all_input_names,
                 thre = thre_in,
                 plotSingleCNEEs = FALSE,
                 SIMPLIFY = F)

## get cal into a table
Ncol = length(all_input)
nectars_clades_out_tb = lapply(1:Ncol, function(i){
  cladeouts[[i]][[1]] %>% mutate('ori' = all_input_names[[i]] )
}) %>% bind_rows() 
dim(nectars_clades_out_tb) # 1451508       6

# ggplot(nectars_clades_out_tb)+
#   geom_histogram(aes(x = exp)) +
#   facet_grid(.~ori)

# cal core non-nectar clades (per clade) ====
all_input_con = list(swift_input, falcon_input, lyrebird_input, passeri1_input)
all_input_con_names = list('swifts', 'falcons', 'lyrebirds', 'passerides')
thre_in = 0.9
cladeouts_con = mapply(clade_specific,
                 clade = all_input_con,
                 name_of_clade = all_input_con_names,
                 thre = thre_in,
                 plotSingleCNEEs = FALSE,
                 SIMPLIFY = F)

## get cal into a table
Ncol = length(all_input_con)
nonnectars_clades_out_tb = lapply(1:Ncol, function(i){
  cladeouts_con[[i]][[1]] %>% mutate('ori' = all_input_con_names[[i]] )
}) %>% bind_rows() 
dim(nonnectars_clades_out_tb) # 1451508       6

path = paste0(project_path, "exp_all_nectar_correctExp_", Sys.Date(), ".csv")
write_csv(nectars_clades_out_tb, file = path, na = "NA")

path = paste0(project_path, "exp_all_control_correctExp_", Sys.Date(), ".csv")
write_csv(nonnectars_clades_out_tb, file = path, na = "NA")
```


## loose set [5]
```{r}
all_branches = c(tree$tip.label, tree$node.label)
background_branches = setdiff(all_branches, ambiguous) # 72
# core_nectars

thre_in = 0.9
bf1_cutoff = 10

nectar_tmp0 = nectars_clades_out_tb %>% 
  filter(logbf1 > bf1_cutoff) 

nectar_pp_count = postZpl %>% 
    filter(`Locus ID` %in% nectar_tmp0$`Locus ID`) %>%
    filter(label %in% core_nectars) %>% 
    group_by(`Locus ID`) %>% 
    summarise(count_nect = sum(pp>thre_in)) 
dim(nectar_pp_count)

bg_pp_count = postZpl %>% 
    filter(`Locus ID` %in% nectar_tmp0$`Locus ID`) %>%
    filter(label %in% background_branches) %>% 
    group_by(`Locus ID`) %>% 
    summarise(count_bg = sum(pp>thre_in)) 
dim(bg_pp_count)

pp_count = nectar_pp_count %>% 
  left_join(bg_pp_count) %>% 
  mutate(relevance = count_nect/count_bg) 

loose_set0 = pp_count %>% 
  filter(relevance > 0.5) %>%
    left_join(nectars_clades_out_tb) %>%
    arrange(desc(relevance))

nectar_clade_order = c("hummingbirds", "parrots", "honeyeaters", "sunbirds")
nectar_clade_order_union = c("union", nectar_clade_order)

loose_set = lapply(1:4, function(i){
  # i=1
  nectar_tmp = loose_set0 %>% 
    filter(ori == nectar_clade_order[i]) 
  return(nectar_tmp)
}) %>% bind_rows()

n_union = n_distinct(loose_set$`Locus ID`) # 3297 <- union

p_loose = loose_set %>% 
  group_by(ori) %>% 
  summarise(count = n_distinct(`Locus ID`)) %>%
  bind_rows(tibble(ori = "union", count = n_union)) %>% 
  mutate(ori = fct_relevel(ori, nectar_clade_order_union)) %>% 
  ggplot(aes(x = ori, y = count)) +
  geom_col(fill = 'steelblue') + 
  geom_text(aes(label = count), vjust = 0, size = 3) +
  scale_y_continuous(expand = c(0,0,0.05, 0)) + theme_m +
  theme(axis.title.x = element_blank()); p_loose
```

#### CNEE look up
```{r fig.height=4*3, fig.width=10 }
nectar_clade_order = c("hummingbirds", "parrots", "honeyeaters", "sunbirds")

p_tmp = 
  mapply(
    clade = nectar_clade_order, sample_n = 5, 
    
    FUN = function(clade, sample_n){
      # clade = nectar_clade_order
      # sample_n = 5
      input = loose_set %>% filter(ori == clade)
      cneeIDs = sample(input$`Locus ID`, sample_n)
      pl = lapply(cneeIDs, FUN = plot_single_cnee, target_species = targets, 
                  show_legend = FALSE, show_all_tips = FALSE)
      pl2 = cowplot::plot_grid(plotlist = pl, ncol = 5) + 
        cowplot::draw_label(paste0(clade), x = 0, y = .5, vjust = 1, hjust = 0); 
    }, SIMPLIFY = FALSE)

cowplot::plot_grid(plotlist = p_tmp, ncol = 1)

# x = postZpl %>% filter(`Locus ID` == '2585-54') %>% filter(label %in% unlist(hummingbird_input))
# x = loose_set  %>% filter(`Locus ID` == '2585-54')
# x = nectars_clades_out_tb %>% filter(`Locus ID` == '1117-79') 
# y = nonnectars_clades_out_tb %>% filter(`Locus ID` == '1117-79')
```


## narrow set
```{r}
bf2_cutoff = 3
narrow_set3 = loose_set %>% filter(logbf2 > bf2_cutoff)

n_union = n_distinct(narrow_set3$`Locus ID`) # 3297 <- union

p_narrow3 = narrow_set3 %>% 
  group_by(ori) %>% 
  summarise(count = n_distinct(`Locus ID`)) %>%
  bind_rows(tibble(ori = "union", count = n_union)) %>% 
  mutate(ori = fct_relevel(ori, nectar_clade_order_union)) %>% 
  ggplot(aes(x = ori, y = count)) +
  geom_col(fill = 'steelblue') + 
  geom_text(aes(label = count), vjust = 0, size = 3) +
  scale_y_continuous(expand = c(0,0,0.05, 0)) + theme_m +
  theme(axis.title.x = element_blank()); p_narrow3
```

#### CNEE look up - narrow_set3 only
```{r fig.height=4*3, fig.width=10 }
nectar_clade_order = c("hummingbirds", "parrots", "honeyeaters", "sunbirds")

p_tmp = 
  mapply(
    clade = nectar_clade_order, sample_n = 5, 
    
    FUN = function(clade, sample_n){
      # clade = nectar_clade_order
      # sample_n = 5
      input = narrow_set3  %>% filter(ori == clade) 
      cneeIDs = sample(input$`Locus ID`, sample_n)
      pl = lapply(cneeIDs, FUN = plot_single_cnee, target_species = targets, 
                  show_legend = FALSE, show_all_tips = FALSE)
      pl2 = cowplot::plot_grid(plotlist = pl, ncol = 5) + 
        cowplot::draw_label(paste0(clade), x = 0, y = .5, vjust = 1, hjust = 0); 
    }, SIMPLIFY = FALSE)

cowplot::plot_grid(plotlist = p_tmp, ncol = 1)

# x = postZpl %>% filter(`Locus ID` == '1117-79') %>% filter(label %in% unlist(hummingbird_input))
# x = nectars_clades_out_tb %>% filter(`Locus ID` == '1117-79') 
# y = nonnectars_clades_out_tb %>% filter(`Locus ID` == '1117-79')
```

#### CNEE look up
```{r fig.height=4*3, fig.width=10 }
nectar_clade_order = c("hummingbirds", "parrots", "honeyeaters", "sunbirds")

p_tmp = 
  mapply(
    clade = nectar_clade_order, sample_n = 5, 
    FUN = function(clade, sample_n){
      # clade = nectar_clade_order[1]
      # sample_n = 5
      tmp_broad  = specific_set2 %>% filter(ori == clade) 
      tmp_broad2 = specific_set22 %>% filter(ori == clade) 
      input = tmp_broad %>% filter(!`Locus ID` %in% tmp_broad2$`Locus ID`) 
      cneeIDs = sample(input$`Locus ID`, sample_n)
      pl = lapply(cneeIDs, FUN = plot_single_cnee, target_species = targets, 
                  show_legend = FALSE, show_all_tips = FALSE)
      pl2 = cowplot::plot_grid(plotlist = pl, ncol = 5) + 
        cowplot::draw_label(paste0(clade), x = 0, y = .5, vjust = 1, hjust = 0); 
    }, SIMPLIFY = FALSE)

cowplot::plot_grid(plotlist = p_tmp, ncol = 1)

```

## output nectar-acc bed
### loose
```{r output_NectarSpecificSet}

RUN = "Loose" # remember to change input
lapply(tmp_sets, function(s){
  # s = "hummingbirds"
  input = loose_set
  tmp_set_cnee = input %>% filter(ori == s) %>%
    left_join(score_ids_sel, by = c("Locus ID" = "phyloacc.id")) %>%
    dplyr::select(id) %>% pull()
  bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
  
  # t = gsub("\\+", "_", s)
  path = paste0(out_path, RUN, "_", s, ".bed")
  write_tsv(bed, path, col_names = F)
})

path = paste0(out_path, RUN, "_", "union", ".bed")
tmp_set_cnee = loose_set %>% 
    left_join(score_ids_sel, by = c("Locus ID" = "phyloacc.id")) %>%
    dplyr::select(id) %>% pull()
bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
write_tsv(bed, path, col_names = F)

# convergent - cnee lv
specific_set_con = loose_set %>% 
  group_by(`Locus ID`) %>% 
  mutate(n_clades = n_distinct(ori))
max_nway = max(unique(specific_set_con$n_clades))

lapply(2:max_nway, function(i){
  input = specific_set_con
  tmp_set_cnee = input %>% filter(n_clades == i) %>%
    left_join(score_ids_sel, by = c("Locus ID" = "phyloacc.id")) %>%
    dplyr::select(id) %>% pull()
  bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
  
  # t = gsub("\\+", "_", s)
  path = paste0(out_path, RUN, "_congvergent", i, "way.bed")
  write_tsv(bed, path, col_names = F)
})

# convergent >=2 - cnee lv
tmp_set_cnee = specific_set_con %>% filter(n_clades >= 2) %>%
    left_join(score_ids_sel, by = c("Locus ID" = "phyloacc.id")) %>%
    dplyr::select(id) %>% pull()
bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
n_distinct(bed$id) # 1570

path = paste0(out_path, RUN, "_congvergent_gteq2way.bed")
write_tsv(bed, path, col_names = F)

# convergent >=3 
tmp_set_cnee = specific_set_con %>% filter(n_clades >= 3) %>%
    left_join(score_ids_sel, by = c("Locus ID" = "phyloacc.id")) %>%
    dplyr::select(id) %>% pull()
bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
n_distinct(bed$id) # 495
  
path = paste0(out_path, RUN, "_congvergent_gteq3way.bed")
write_tsv(bed, path, col_names = F)

# full specific
path = paste0(project_path, RUN, "_full_output_", Sys.Date(), "_correctExp.csv")
write_csv(loose_set, file = path, na = "NA")

# convergent - gene lv
path = "/home/mpg08/mko/Nectar/analysis/CNEEanalysis_2021/03_phyloacc/2024ver/nectar_nodollo/great/overlap/cnee_gene_assignment_basalPlusExtension_proteincoding_allcnees.bed"
allcnees = read_tsv(path, col_names = c("chr", "start", "end", "id", "chr2", "start2", "end2", "gene")) %>% 
      mutate(gene = gsub("LOC107051846", "IRS1", gene),
           gene = gsub("LOC107055431", "NEDD4L", gene),
           gene = gsub("gga-mir-6690", "CTSB", gene),
           gene = gsub("gga-mir-12216", "B2M", gene),
           gene = gsub("gga-mir-12230", "SLC27A4", gene),
           gene = gsub("^Ii$", "CD74", gene)) 

path = "/home/mpg08/mko/Nectar/analysis/CNEEanalysis_2021/03_phyloacc/2024ver/nectar_nodollo/great/overlap/Loose/"
target_file = list.files(path, ".*_basalPlusExtension_\\w+.bed", full.names = T); target_file # great default assign rule

acccnees_100kb_withcnees0_loose = lapply(target_file, function(s){
  # s = target_file[1]
  tmp_name = basename(s)
  tmp_name = gsub("cnee_gene_assignment_basalPlusExtension_proteincoding_Loose_", "", tmp_name)
  tmp_name = gsub(".bed", "", tmp_name)
  print(paste0("current: ", tmp_name))
  x = read_tsv(s, col_names = c("chr", "start", "end", "id", "chr2", "start2", "end2", "gene")) %>% 
    dplyr::select(gene, id) %>%
    bind_cols('group' = tmp_name) 
}) %>% bind_rows()  %>% 
  # mutate(group = gsub("_2", "", group)) %>% 
  mutate(group = gsub("sunbirds_flowerpecker", "sunbirds", group)) %>% 
  mutate(gene = gsub("LOC107051846", "IRS1", gene),
         gene = gsub("LOC107055431", "NEDD4L", gene),
         gene = gsub("gga-mir-6690", "CTSB", gene),
         gene = gsub("gga-mir-12216", "B2M", gene),
         gene = gsub("gga-mir-12230", "SLC27A4", gene),
         gene = gsub("^Ii$", "CD74", gene))
head(acccnees_100kb_withcnees0_loose)

acccnees_100kb_withcnees_loose = acccnees_100kb_withcnees0_loose %>% 
  mutate(set = paste0("acc_", group), 
         val = 1,
         analysis = "acc") %>% 
  dplyr::select(set, gene, val, analysis, group) %>% 
  distinct()
dim(acccnees_100kb_withcnees_loose) # 25585    5
head(acccnees_100kb_withcnees_loose)
# set                gene     val analysis group    
# acc_2wayconvergent KLHL1      1 acc      congvergent2way
# acc_2wayconvergent PLXNA4     1 acc      congvergent2way
unique(acccnees_100kb_withcnees_loose$group)

## keep only sets of 4 cloades
acccnees_100kb_withcnees0_loose_sel = acccnees_100kb_withcnees0_loose %>%
  filter(group %in% c("honeyeaters", "hummingbirds", "parrots", "sunbirds" )) %>% 
  dplyr::rename(clade = group)

## make convergent table (gene lv)
genelv_conv_b = acccnees_100kb_withcnees0_loose_sel %>% 
  group_by(gene) %>% 
  mutate(n_clade = length(unique(clade))) %>% 
  dplyr::select(gene, n_clade) %>% 
  distinct()
table(genelv_conv_b$n_clade)

## at least 2 clades
genelv_conv_b_gteq2 = genelv_conv_b %>% filter(n_clade>=2)
n_distinct(genelv_conv_b_gteq2$gene) # 3681

path = paste0(out_path, RUN, "_congvergent_genelv_gteq2way_genelist.tsv")
write_tsv(genelv_conv_b_gteq2, path)

### get cnees
genelv_conv_b_gteq2_cnee = genelv_conv_b_gteq2 %>% 
  left_join(allcnees) %>% 
  # keep only cnees in the nectar set
  filter(id %in% acccnees_100kb_withcnees0_loose_sel$id)
n_distinct(genelv_conv_b_gteq2_cnee$id) # 10991

tmp_set_cnee = genelv_conv_b_gteq2_cnee %>%
    dplyr::select(id) %>% distinct() %>% pull()
bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
path = paste0(out_path, RUN, "_congvergent_genelv_gteq2way.bed")
write_tsv(bed, path, col_names = F)

## at least 3 clades
genelv_conv_b_gteq2 = genelv_conv_b %>% filter(n_clade>=3)
n_distinct(genelv_conv_b_gteq2$gene) # 3681

path = paste0(out_path, RUN, "_congvergent_genelv_gteq3way_genelist.tsv")
write_tsv(genelv_conv_b_gteq2, path)

### get cnees
genelv_conv_b_gteq2_cnee = genelv_conv_b_gteq2 %>% 
  left_join(allcnees) %>% 
  # keep only cnees in the nectar set
  filter(id %in% acccnees_100kb_withcnees0_loose_sel$id)
n_distinct(genelv_conv_b_gteq2_cnee$id) # 10991

tmp_set_cnee = genelv_conv_b_gteq2_cnee %>%
    dplyr::select(id) %>% distinct() %>% pull()
bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
path = paste0(out_path, RUN, "_congvergent_genelv_gteq3way.bed")
write_tsv(bed, path, col_names = F)

## convergent exactly n clade
max_nway = max(unique(specific_set_con$n_clades))

lapply(2:max_nway, function(i){
  input = genelv_conv_b %>% filter(n_clade ==i) 
  path = paste0(out_path, RUN, "_congvergent_genelv_", i, "way_genelist.tsv")
  write_tsv(input, path)

  ### get cnees
  input_cnee = input %>% 
    left_join(allcnees) %>% 
    # keep only cnees in the nectar set
    filter(id %in% acccnees_100kb_withcnees0_loose_sel$id)
  n_distinct(input_cnee$id) # 10991
  
  tmp_set_cnee = input_cnee %>%
      dplyr::select(id) %>% distinct() %>% pull()
  bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
  path = paste0(out_path, RUN, "_congvergent_genelv_", i, "way.bed")
  write_tsv(bed, path, col_names = F)
})

```

### logBF3 > 10
```{r}
sum(score$logbf3>10) # 58363
x = score %>% filter(logbf3>10)

tmp_set_cnee = x %>% 
    left_join(score_ids_sel, by = c("phyloacc.id")) %>%
    dplyr::select(id) %>% pull()
bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)

path = paste0(out_path, "logBFgt10", "_", ".bed")
write_tsv(bed, path, col_names = F)

```


### narrow logBF > 3
```{r output_NectarSpecificSet}

RUN = "NarrowBF2_gt3" # remember to change input
lapply(tmp_sets, function(s){
  # s = "hummingbirds"
  input = narrow_set3
  tmp_set_cnee = input %>% filter(ori == s) %>%
    left_join(score_ids_sel, by = c("Locus ID" = "phyloacc.id")) %>%
    dplyr::select(id) %>% pull()
  bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
  
  # t = gsub("\\+", "_", s)
  path = paste0(out_path, RUN, "_", s, ".bed")
  write_tsv(bed, path, col_names = F)
})

path = paste0(out_path, RUN, "_", "union", ".bed")
tmp_set_cnee = narrow_set3 %>% 
    left_join(score_ids_sel, by = c("Locus ID" = "phyloacc.id")) %>%
    dplyr::select(id) %>% pull()
bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
write_tsv(bed, path, col_names = F)

# convergent
specific_set_con = narrow_set3 %>% 
  group_by(`Locus ID`) %>% 
  mutate(n_clades = n_distinct(ori))
max_nway = max(unique(specific_set_con$n_clades))

lapply(2:max_nway, function(i){
  input = specific_set_con
  tmp_set_cnee = input %>% filter(n_clades == i) %>%
    left_join(score_ids_sel, by = c("Locus ID" = "phyloacc.id")) %>%
    dplyr::select(id) %>% pull()
  bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
  
  # t = gsub("\\+", "_", s)
  path = paste0(out_path, RUN, "_congvergent", i, "way.bed")
  write_tsv(bed, path, col_names = F)
})

# convergent >=2 
tmp_set_cnee = specific_set_con %>% filter(n_clades >= 2) %>%
    left_join(score_ids_sel, by = c("Locus ID" = "phyloacc.id")) %>%
    dplyr::select(id) %>% pull()
bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
  
path = paste0(out_path, RUN, "_congvergent_gteq2way.bed")
write_tsv(bed, path, col_names = F)

# convergent >=3 
tmp_set_cnee = specific_set_con %>% filter(n_clades >= 3) %>%
    left_join(score_ids_sel, by = c("Locus ID" = "phyloacc.id")) %>%
    dplyr::select(id) %>% pull()
bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
  
path = paste0(out_path, RUN, "_congvergent_gteq3way.bed")
write_tsv(bed, path, col_names = F)
  

# full specific
path = paste0(project_path, RUN, "_full_output_", Sys.Date(), "_correctExp.csv")
write_csv(narrow_set3, file = path, na = "NA")


# convergent - gene lv
path = "/home/mpg08/mko/Nectar/analysis/CNEEanalysis_2021/03_phyloacc/2024ver/nectar_nodollo/great/overlap/cnee_gene_assignment_basalPlusExtension_proteincoding_allcnees.bed"
allcnees = read_tsv(path, col_names = c("chr", "start", "end", "id", "chr2", "start2", "end2", "gene")) %>% 
      mutate(gene = gsub("LOC107051846", "IRS1", gene),
           gene = gsub("LOC107055431", "NEDD4L", gene),
           gene = gsub("gga-mir-6690", "CTSB", gene),
           gene = gsub("gga-mir-12216", "B2M", gene),
           gene = gsub("gga-mir-12230", "SLC27A4", gene),
           gene = gsub("^Ii$", "CD74", gene)) 

path = "/home/mpg08/mko/Nectar/analysis/CNEEanalysis_2021/03_phyloacc/2024ver/nectar_nodollo/great/overlap/NarrowBF2_gt3/"
target_file = list.files(path, ".*_basalPlusExtension_\\w+.bed", full.names = T); target_file # great default assign rule

acccnees_100kb_withcnees0_narrow3 = lapply(target_file, function(s){
  # s = target_file[1]
  tmp_name = basename(s)
  tmp_name = gsub("cnee_gene_assignment_basalPlusExtension_proteincoding_NarrowBF2_gt3_", "", tmp_name)
  tmp_name = gsub(".bed", "", tmp_name)
  print(paste0("current: ", tmp_name))
  x = read_tsv(s, col_names = c("chr", "start", "end", "id", "chr2", "start2", "end2", "gene")) %>% 
    dplyr::select(gene, id) %>%
    bind_cols('group' = tmp_name) 
}) %>% bind_rows()  %>% 
  # mutate(group = gsub("_2", "", group)) %>% 
  mutate(group = gsub("sunbirds_flowerpecker", "sunbirds", group)) %>% 
  mutate(gene = gsub("LOC107051846", "IRS1", gene),
         gene = gsub("LOC107055431", "NEDD4L", gene),
         gene = gsub("gga-mir-6690", "CTSB", gene),
         gene = gsub("gga-mir-12216", "B2M", gene),
         gene = gsub("gga-mir-12230", "SLC27A4", gene),
         gene = gsub("^Ii$", "CD74", gene))
head(acccnees_100kb_withcnees0_narrow3)

acccnees_100kb_withcnees_narrow3 = acccnees_100kb_withcnees0_narrow3 %>% 
  mutate(set = paste0("acc_", group), 
         val = 1,
         analysis = "acc") %>% 
  dplyr::select(set, gene, val, analysis, group) %>% 
  distinct()
dim(acccnees_100kb_withcnees_narrow3) # 7559    5
head(acccnees_100kb_withcnees_narrow3)
# set                gene     val analysis group    
# acc_2wayconvergent KLHL1      1 acc      congvergent2way
# acc_2wayconvergent PLXNA4     1 acc      congvergent2way
unique(acccnees_100kb_withcnees_narrow3$group)

## keep only sets of 4 cloades
acccnees_100kb_withcnees0_narrow3_sel = acccnees_100kb_withcnees0_narrow3 %>%
  filter(group %in% c("honeyeaters", "hummingbirds", "parrots", "sunbirds" )) %>% 
  dplyr::rename(clade = group)

## make convergent table (gene lv)
genelv_conv_b = acccnees_100kb_withcnees0_narrow3_sel %>% 
  group_by(gene) %>% 
  mutate(n_clade = length(unique(clade))) %>% 
  dplyr::select(gene, n_clade) %>% 
  distinct()
table(genelv_conv_b$n_clade)

## at least 2 clades
genelv_conv_b_gteq2 = genelv_conv_b %>% filter(n_clade>=2)
n_distinct(genelv_conv_b_gteq2$gene) # 1877

path = paste0(out_path, RUN, "_congvergent_genelv_gteq2way_genelist.tsv")
write_tsv(genelv_conv_b_gteq2, path)

### get cnees
genelv_conv_b_gteq2_cnee = genelv_conv_b_gteq2 %>% 
  left_join(allcnees) %>% 
  # keep only cnees in the nectar set
  filter(id %in% acccnees_100kb_withcnees0_narrow3_sel$id)
n_distinct(genelv_conv_b_gteq2_cnee$id) # 2309

tmp_set_cnee = genelv_conv_b_gteq2_cnee %>%
    dplyr::select(id) %>% distinct() %>% pull()
bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
path = paste0(out_path, RUN, "_congvergent_genelv_gteq2way.bed")
write_tsv(bed, path, col_names = F)

## at least 3 clades
genelv_conv_b_gteq2 = genelv_conv_b %>% filter(n_clade>=3)
n_distinct(genelv_conv_b_gteq2$gene) # 3681

path = paste0(out_path, RUN, "_congvergent_genelv_gteq3way_genelist.tsv")
write_tsv(genelv_conv_b_gteq2, path)

### get cnees
genelv_conv_b_gteq2_cnee = genelv_conv_b_gteq2 %>% 
  left_join(allcnees) %>% 
  # keep only cnees in the nectar set
  filter(id %in% acccnees_100kb_withcnees0_loose_sel$id)
n_distinct(genelv_conv_b_gteq2_cnee$id) # 10991

tmp_set_cnee = genelv_conv_b_gteq2_cnee %>%
    dplyr::select(id) %>% distinct() %>% pull()
bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
path = paste0(out_path, RUN, "_congvergent_genelv_gteq3way.bed")
write_tsv(bed, path, col_names = F)

## convergent exactly n clade
max_nway = max(unique(specific_set_con$n_clades))

lapply(2:max_nway, function(i){
  input = genelv_conv_b %>% filter(n_clade ==i) 
  path = paste0(out_path, RUN, "_congvergent_genelv_", i, "way_genelist.tsv")
  write_tsv(input, path)

  ### get cnees
  input_cnee = input %>% 
    left_join(allcnees) %>% 
    # keep only cnees in the nectar set
    filter(id %in% acccnees_100kb_withcnees0_narrow3_sel$id)
  n_distinct(input_cnee$id) # 2:882;3:846;4:969
  
  tmp_set_cnee = input_cnee %>%
      dplyr::select(id) %>% distinct() %>% pull()
  bed = ggal_final_cnees %>% filter(id %in% tmp_set_cnee)
  path = paste0(out_path, RUN, "_congvergent_genelv_", i, "way.bed")
  write_tsv(bed, path, col_names = F)
})


```



## output setup
```{r}

Factor_mmtoin = 0.0393701
Width_HalfCol = 85*Factor_mmtoin
AxisTxFontSizeSize = 6

```



## MAT2B CNEE302671
```{r}

gene_id = "MAT2B"
cneeIDs = "CNEE302671"
cneeIDs_new = score_ids_sel %>% filter(id %in% cneeIDs) %>% pull( phyloacc.id)

pl = lapply(cneeIDs_new, FUN = plot_single_cnee, target_species = targets, show_legend = TRUE, show_all_tips = TRUE, xratio = 0.8)
p = cowplot::plot_grid(plotlist = pl, ncol = length(cneeIDs)); p

graph_path = paste0(getwd(), "/", gene_id, "_", cneeIDs, "_", Sys.Date(),".pdf"); graph_path
pdf(graph_path, width = Width_HalfCol*1*1, height = Width_HalfCol*1.1*1, pointsize = AxisTxFontSizeSize)
p
dev.off()

# no lab
pl = lapply(cneeIDs_new, FUN = plot_single_cnee, target_species = targets, 
            show_legend = TRUE, show_all_tips = FALSE, xratio = 0.8)
p = cowplot::plot_grid(plotlist = pl, ncol = length(cneeIDs)); p

graph_path = paste0(getwd(), "/", gene_id, "_", cneeIDs, "_nolab_", Sys.Date(),".pdf"); graph_path
pdf(graph_path, width = Width_HalfCol*1*1, height = Width_HalfCol*1.1*1, pointsize = AxisTxFontSizeSize)
p
dev.off()
```


